---
title: "New Book Program"
author: "Maenya"
date: "2025-02-18"
output: html_document
---

# Introduction
The role of the analysis is to evaluate the impact of a new program launched on July 1, 2019 to encourage customers to buy more books. 

### Research Question: 
Did the new program launched on July 1, 2019 to encourage customers to buy more book?

### Loading the necessary libraries
```{r}
library(tidyverse)
library(lubridate)
```

### Loading the data
```{r}
book <- read.csv("sales2019.csv")
```

### Understanding the data
```{r}
head(book)
str(book)
glimpse(book)
```

### checking missing data

```{r}
#checking missing data
missing_percentage <- colSums(is.na(book)) / nrow(book) * 100
print(missing_percentage)
```

#### We have significant missing data
#### We will do away with them inorder to get a good analysis report and workout
```{r}
# a. Changing to date format

book1 <- book %>%
  select(date, total_purchased, customer_type, user_submitted_review) %>%
  mutate(date = as.Date(date, format = "%m/%d/%Y"))
```

### converting to M/Y to help in summing monthly sales

```{r}
book2 <- book1 %>%
  mutate(month = format(date, "%Y-%m"))

str(book2$month)
glimpse(book2)
print(book2)
  
```  
#### Summing the sales per month
```{r}
 book3<- book2 %>%
  group_by(month)%>%
  summarise(purchased=sum(total_purchased,na.rm = TRUE))
   
 print(book3)
```


#### checking the trend
```{r} 
 ggplot(book3, aes(x = factor(month, levels = unique(month)), y = purchased)) +
   geom_col(fill = "blue") +
   coord_flip() +
   labs(title = "Total Sales by Month",
        x = "Month",
        y = "Total Sales") +
   theme_minimal()
```

### Line Graph
```{r}
 ggplot(book3, aes(x = as.Date(paste0(month, "-01")), y = purchased, group = 1)) +
   geom_line(color = "blue", size = 1) +  # Line color and thickness
   geom_point(color = "red", size = 2) +  # Add points for clarity
   labs(title = "Total Sales Over Time",
        x = "Month",
        y = "Total Sales") +
   theme_minimal() +
   scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +  # Format x-axis
   theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels for readability
 
```
 
### Confirming the mean of sales before and after July
#### Before that, we should change the date format to MM/YY
 
```{r}
book4 <- book3 %>%
   mutate(month = as.Date(paste0("20", substr(month, 3, 7), "-01"))) 
 str(book4)
``` 
 
####Applying After and Before July
```{r}
 book5 <- book4 %>%
   mutate(period = ifelse(month < as.Date("2019-07-01"), "Before", "After"))
 print(book5)
```

### Summarizing the descriptive statistics for BEFORE and AFTER
```{r}
summary_stats <- book5 %>%
   group_by(period) %>%
   summarise(
     total_sales = sum(purchased, na.rm = TRUE),
     avg_sales = mean(purchased, na.rm = TRUE),
     sd_sales = sd(purchased, na.rm = TRUE),
     median_sales = median(purchased, na.rm = TRUE),
     min_sales = min(purchased, na.rm = TRUE),
     max_sales = max(purchased, na.rm = TRUE),
     count = n()
   )
 
 print(summary_stats)
```
 
### Line Graph after descritive stat
```{r}
 library(ggplot2)
 
 ggplot(book5, aes(x = month, y = purchased, group = 1)) +
   geom_line(color = "blue", size = 1) +  # Blue line for sales trend
   geom_point(aes(color = period), size = 3) +  # Different colors for Before/After
   geom_vline(xintercept = as.Date("2019-07-01"), linetype = "dashed", color = "red") +  # Mark July 2019
   labs(title = "Sales Trends Before and After the New Book Program",
        x = "Month",
        y = "Total Sales") +
   theme_minimal() +
   scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +  # Show all months
   scale_color_manual(values = c("Before" = "black", "After" = "red")) +  # Colors for periods
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

``` 
###3. Independent t-Test (Compare Mean Sales Before vs. After July)
```{r}
 t_test_result <- t.test(purchased ~ period, data = book5, var.equal = TRUE)  
 
 print(t_test_result) 
 
 #p-value = 0.6367 → This is much greater than 0.05, meaning we fail to reject the null hypothesis. There is no significant difference in sales before and after July.
 #Mean sales (Before) = 1440.33, Mean sales (After) = 1420.83 → Sales actually decreased slightly after July.
 # 95% Confidence Interval: (-108.7, 69.7) → This range includes 0, meaning there is no clear increase or decrease in sales due to the new book program.
 
 #4. Regression Analysis (Account for Other Factors)
 # Convert month to numerical format for regression
 book6 <- book5 %>%
   mutate(month_numeric = as.numeric(format(month, "%m")),  # Extract month as a number
          period_binary = ifelse(period == "After", 1, 0))  # Convert period to binary (0=Before, 1=After)
``` 
### Run the regression model
```{r}
 model <- lm(purchased ~ period, data = book6)
 
 # Display summary of the model
 summary(model)
```
 
 
# Conclusion
a. The new book program did not significantly impact sales.
b. Sales were relatively stable before and after July, with only minor variation.
c. Other factors likely influence sales more than the program itself.
